\chapter{General Relativity}



\section{Preliminaries}


\subsection{What is flat space}

Space is flat if there exists a way to chose the metric tensor, such that it's the knonecker delta. $\eta^{x,y} = \delta_{x,y}$.

\begin{small}
  A piece of flat paper that is folded is still flat, it has no curvature, it can be made flat again and the notion of distance is still the same as when it was flat. This is just extrinsic curvature, which comes from the way it's embedded in a higher dimentional space.
\end{small}


\subsection{Tensor Analysis}

\paragraph{} When we take the components as scalar that sum to it, the components are contravarient components.
\begin{equation}
  V = V^1 \hat{e}_1 + V^2 \hat{e}_2 + V^3 \hat{e}_3
\end{equation}

And if we find the components using dot products, we get the covarient components.
\begin{equation}
  V_n = V \cdot \hat{e}_n = \Sigma_n V^n (e_n \cdot e_m)
\end{equation}

These are exactly the same thing when we are in the Cartesian coordinates.

\subsubsection{Transformation Rules of Tensors}

\paragraph{} For contravarient index:
\begin{equation}
  dy^m = \frac{\partial y^m}{\partial x^n} dx^n
\end{equation}
\begin{equation}
  \frac{\partial S}{\partial y^m} = \frac{\partial x^n}{\partial y^m} \frac{\partial S}{\partial x^n}
\end{equation}

\paragraph{} For covarient index:
\begin{equation}
  dy^m = \frac{\partial y^m}{\partial x^n} dx^n
\end{equation}
\begin{equation}
  \frac{\partial S}{\partial y^m} = \frac{\partial x^n}{\partial y^m} \frac{\partial S}{\partial x^n}
\end{equation}

\paragraph{} For Mixed tensors (Rank 2):
\begin{equation}
  (W^\prime)^m_n = \frac{\partial y^m}{\partial x^p} \frac{\partial x^q}{\partial y^n} W^p_q
\end{equation}
\paragraph{} For Covarient tensors (Rank 2):
\begin{equation}
  (W^\prime)_{mn} = \frac{\partial x^p}{\partial y^m} \frac{\partial x^q}{\partial y^n} W_{pq}
\end{equation}
\paragraph{} For contravarient tensors (Rank 2):
\begin{equation}
  (W^\prime)^{mn} = \frac{\partial y^m}{\partial x^p} \frac{\partial y^n}{\partial x^q} W^{pq}
\end{equation}

\subsubsection{Addition, Multiplication, Contraction}

\paragraph{} We only add tensor if they have indices of the same kind, that is:
\begin{equation}
  T^{m_1,m_2...m_k}_{n_1,n_2...n_l} + S^{m_1,m_2...m_k}_{n_1,n_2...n_l} = (T + S)^{m_1,m_2...m_k}_{n_1,n_2...n_l}
\end{equation}

\paragraph{} We can multiply any two tensors (We do get tensors of higher rank):
\begin{eqnarray}
  V^m \otimes W_n &=& X^m_n  \\
  V^m \otimes W^n &=& X^{mn} \\
  V_m \otimes W_n &=& X_{mn}
\end{eqnarray}

\paragraph{} The generalization of inner product of two vectors to tensors is called it's contraction.
Proof/Intuition of contraction:
\begin{equation}
  (V^m W_m)^\prime = \frac{\partial }{\partial} (V^a W_b)
\end{equation}
\paragraph{} When we contract an upper with a lower index, we reduce the number of indices (rank) by 2, as both of them become dummy indices, and we sum over them.

\subsubsection{The Metric Tensor}

\begin{equation}
  ds^2 = g(x)_{mn} dx^m dx^n
\end{equation}
\paragraph{} The metric tensor is always symetric. Everyone agrees on the length of every vector, though not on the individual components, when viewed from different frames.

Transformation of the metric tensor:
\begin{equation}
  g_{mn}(x) dx^m dx^n = g_{pq}^\prime dy^p dy^q = g_{mn}(x) = \mathbf{g_{pq} \frac{\partial x^m}{\partial x^n} \frac{\partial y^p}{\partial y^q}} dy^p dy^q
\end{equation}

\paragraph{} We have two metric tensors, one with covarient and one with contravarient indices, defined as:
\begin{equation}
  g_{mn} g^{np} = \delta_m^p
\end{equation}
So, one is the inverse of the other.



\section{Cartesian Geometry}


\subsection{Flatness of a Space}

If $g_mn \neq \delta_mn$ (metric tensor), this does not mean that the space is flat. This just means we have bad coordinates. Let's say there is some quantity Riemann $R$.

In Riemannian Geometry, all lengths are positive, all of their squares are also positive. Equivalently, the metric tensor is positive definite, i.e. all it's eigen values are positive.


\subsection{Gaussian Normal Coordinates}

At every point, we can chose Gaussian Normal Coordinates (unique upto all rotations), such that $g_{mn} = \delta_{mn}$ at $r = r_0$. Also we can chose $\frac{\partial g_{mn}}{\partial x^r} = 0$, i.e. our metric can be chosen to have first derivative 0. We can only make higher derivatives 0 iff the space is flat, but not otherwise $\frac{\partial^2 g_{mn}}{\partial x^r \partial x^s} \neq 0$.

This can always be done since $x^m = y^m + c^m_{nr} y^n y^r$ is the most general quadratic, and it contains 40 free constants in 4-spaces, since all of $c^m_{nr}$ are all free. But this only allows us to set the derivatives of all 10 variables to 0 with respect to each of the 4 variables, so we have derivative to first order is 0.

\textbf{The meeaning of Covartient derivatives is that they are all 0 in Gaussian Normal Coordinates, and then making Tensors out of them}.
\begin{eqnarray}
  D_s g_mn &=& \frac{\partial g_mn}{\partial x^s} - \Gamma^{t}_{sm} g_{tn} - \Gamma^{t}_{sn} g_{mt} = 0 \\
  D_m g_sn &=& \frac{\partial g_sn}{\partial x^m} - \Gamma^{t}_{sm} g_{tn} - \Gamma^{t}_{mn} g_{st} = 0 \\
  D_n g_sm &=& \frac{\partial g_sm}{\partial x^n} - \Gamma^{t}_{sn} g_{tm} - \Gamma^{t}_{mn} g_{st} = 0 \\
  \Gamma^{t}_{mn} &=& \frac{1}{2} [\partial_n g_{sm} + \partial_m g_{sn} + \partial_s g_{mn}]
\end{eqnarray}
We got the last equation from (1) + (2) - (3), and so we have the value of the Cristophel symbols. They can be 0 in one frame and non-zero in another, so we go back to the tensor frame after doing the derivatives.



\section{Starting off with Gravitation}

\begin{equation}
  D_M V^n = \frac{\partial V^n}{\partial x^m} + \Gamma_{mn}^{r} V^r
\end{equation}