\chapter{Particle Physics}



\section{Quantum Field Theory}


\subsection{Basic Quantum Mechanics}

The following are true for Light, the second equation for momentum is only valid for a photon.
\begin{eqnarray*}
  E &=& \hbar \omega = h f \\
  p &=& \frac{E}{c} = \frac{h}{\lambda}
\end{eqnarray*}


\subsection{The Quantum Field}

\subsubsection{Getting the Schrodinger equation}

\begin{eqnarray}
  \Psi^\dagger(x, t) &=& \sum_k a^+(k) e^{-ikx} e^{+i\omega(k)t} \\
  \Psi(x, t) &=& \sum_k a^-(k) e^{+ikx} e^{-i\omega(k)t}
\end{eqnarray}
Let's talk about a non relativistic particle. In normal units, $\hbar = 1$, so $\omega = E$ and $k = p$.
\begin{eqnarray}
  \frac{\partial \Psi(x, t)}{\partial t} &=& i\omega \sum_k a^+(k) e^{-ikx} e^{+i\omega(k)t} \\
  \frac{\partial^2 \Psi(x, t)}{\partial x^2} &=& -k^2 \sum_k a^-(k) e^{+ikx} e^{-i\omega(k)t}
\end{eqnarray}
So we get the equation (same in form to Schrodinger)
\begin{equation}
  - \frac{1}{2m} \frac{\partial^2 \Psi}{\partial x^2} = i \frac{\partial \Psi}{\partial t}
\end{equation}
Here $\Psi$ is not the wavefunction, it's creators and annahiators on the field. It can create and destroy particles.


\subsection{Scattering Experiment}

Our model be that a particle can get scattered over any time. Scattering is just annahialation and simultaneous creation. We do this to get the following probability of scattering (Where g is the coupling constant, denoting the probability with which the interaction will happen, since mesons interact a lot with protons, electrons less so, gravitons so much less):
\begin{eqnarray}
  \braket{k_f|g \int_{t=-\infty}^{\infty} dt (\Psi^\dagger(0, t) \Psi(0, t))|k_i}
\end{eqnarray}
Since this is a single number term, it gives us the probability of interaction, 


\subsection{Group Velocity and Phase Velocity}

Group velocity is the real velocity of the wave packet. Phase velocity is the velocity of each individual crest that compose the interference pattern of the wave. 

Phase velocity is the following: $v_{phase} = \frac{\omega}{k}$, it's easy to see from the form of the wave. There is an obvious problem, if we increase $\omega$ by a constant, we are just increasing the global energy by a constant, changing wavefunctions by a global phase, which should change nothing but here it's changing the velocity, which proves that phase velocity is not associated with any physical measureable.

We notice that the position where the interference peak is, when two waves of very close by wavenumbers are interfering is where $sin(k x - \omega(k) t) = sin(k^\prime x - \omega(k^\prime) t)$, therefore $x = \frac{\omega(k^\prime) - \omega(k)}{k^\prime - k}$, so we have the following: $v_{group} = \frac{\partial \omega}{\partial k}$.

Group and Velocity are the same when the mass of the wave is 0 (light).